Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes

### 2020-02-07 ms/SI.rmd
For SI load captions separately and add header and footer to tables

# Set up captions
# - SI figures
s_fig_nums <- captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
s_fig_nums(name = "corr-plot", "Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.")
s_fig_nums(name = "cwsd", "Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia.")

# - SI tables
s_table_nums <- captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")
s_table_nums(name = "species-list", "List of ferns from Moorea, French Polynesia included in this study")
s_table_nums(name = "trait-values", "Trait values of ferns on Moorea, French Polynesia")
s_table_nums(name = "ancova", "Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate")
s_table_nums(name = "phylosig-binary-strict", "Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy")
s_table_nums(name = "corr-evo-strict", "Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy")
s_table_nums(name = "model-fit", "Best-fitting models explaining community diversity metrics by climate and growth habit")

# Make short versions of citation functions
s_figure <- pryr::partial(s_fig_nums, display = "cite")
s_table <- pryr::partial(s_table_nums, display = "cite")
```

@@

Article acceptance date: ADD_DATE_WHEN_ACCEPTED

@@

**`r s_table("species-list")`** List of ferns from Moorea, French Polynesia included in this study

'taxon code' is the name used for that taxon in analyses. 'informal variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them. 'name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = @Murdock2003.

**`r s_table("trait-values")`** Trait values of ferns on Moorea, French Polynesia

SLA, specific leaf area. Continuous measurements shown as mean, standard deviation (sample size). 'taxon code' is the name used for that taxon in analyses (see Table S1). For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI.

```{r}
trait_data_for_si %>%
  write_csv(here::here("results/Table_S2.csv"))
```

**`r s_table("ancova")`** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate

@@

```{r climate-ancova}
# Format and write out table as CSV file
climate_ancova_results %>% 
  mutate(
    p.value = signif(p.value, 2) %>% format.pval(., digits = 1),
    meansq = round_t(meansq, 2),
    term = case_when(
      term == "habit" ~ "Growth habit",
      term == "el" ~ "Elevation",
      term == "habit:el" ~ "Growth habit × elevation",
      TRUE ~ term
    )
  ) %>%
  select(`Response variable` = variable, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value) %>%
  write_csv(here::here("results/Table_S3.csv"))

@@

**`r s_table("phylosig-binary-strict")`** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy

```{r phylosig-binary-strict}
# Format and write out table as CSV file
phylosig_binary_traits_strict %>%
  mutate(D = round_t(D, 2)) %>%
  mutate(trait = factor(
    trait, 
    levels = c("habit", "gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Epiphytic growth", "Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
  arrange(trait) %>%
  select(Trait = trait, `Number of presences` = num_present, `Number of absences` = num_absent, D,
         P_rnd = prob_random, P_bm = prob_brownian) %>%
  write_csv(here::here("results/Table_S4.csv"))
```

**`r s_table("corr-evo-strict")`** Pagel's [-@Pagel1994] test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy

```{r corr-evo-strict}
# Format and write out table as CSV file
correlated_evo_test_strict %>%
  mutate(trait = factor(
    trait, 
    levels = c("gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Gemmae", "Glands", "Hairs", "Morphotype"))) %>%
  arrange(trait) %>%
  select(
    Trait = trait,
    `LL (independent model)` = logL_indep,
    `LL (dependent model)` = logL_dep,
    `Likelihood ratio` = likelihood_ratio,
    `P` = pval
  ) %>%
  write_csv(here::here("results/Table_S5.csv"))
```

**`r s_table("model-fit")`** Best-fitting models explaining community diversity metrics by climate and growth habit

```{r best-fit-mods}
# Format and write out table as CSV file
best_fit_div_models_moran %>%
  # Drop frond length and width (highly correlated with rhizome dia.)
  filter(resp_var %in% c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna")) %>%
  mutate(resp_var = factor(
        resp_var, 
        levels = c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna"))) %>%
  mutate(
    resp_var = lvls_revalue(
        resp_var, 
        new_levels = c("Richness", "MPDphy", "MNTDphy", 
                       "MPDfunc", "MNTDfunc",
                       "Dissection", "SLA", "Stipe length", "Rhizome diam.", "Pinna no.")) 
  ) %>%
  arrange(resp_var, delta_AICc) %>%
  mutate(
    modname = modname %>%
      str_replace_all("\\+", " \\+ ") %>%
      str_replace_all("\\.by\\.", " x ") %>%
      str_replace_all("habit", "Growth habit") %>%
      str_replace_all("temp_mean", "Mean temp.") %>%
      str_replace_all("temp_sd", "SD temp.") %>%
      str_replace_all("temp_max", "Max. temp.") %>%
      str_replace_all("vpd_mean", "Mean VPD") %>%
      str_replace_all("vpd_sd", "SD VPD") %>%
      str_replace_all("vpd_min", "Min. VPD")
  ) %>% 
  select(
    `Metric` = resp_var,
    `Model` = modname,
    AICc,
    `R2` = r2,
    delta_AICc,
    `Moran's I` = morans_I,
    `P_moran` = morans_I_pval
  ) %>%
  write_csv(here::here("results/Table_S6.csv"))
```

@@

```{r Fig-S1, fig.height = 8, fig.width = 4, fig.cap = ""}

>>>>

# Load captions
source(here::here("ms/captions.R"))

# Set up table headers
si_table_header <- c(
  "New Phytologist Supporting Information",
  "Authors: Joel H. Nitta, James E. Watkins, Jr., and Charles C. Davis" %>% wrap_quotes,
  "Article acceptance date: 5 February 2020"
)
```

<!-- Set left justification  -->
\raggedright

<!-- Don't indent -->
\setlength{\parindent}{0ex}

@@

Article acceptance date: 5 February 2020

@@

**`r s_table("species-list")`** `r s_table_cap("species-list")` (see separate file)

'taxon code' is the name used for that taxon in analyses. 'informal variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them. 'name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith (2003). 'NA' indicates not applicable.

```{r species-list}
# Set header and footer
header <- c(
  si_table_header,
  s_table_nums("species-list") %>% wrap_quotes
)

# Wrap footer at ca. 100 characters per line. Each line needs to be an element of the
# character vector.
footer <- "'taxon code' is the name used for that taxon in analyses. 'informal variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them. 'name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith (2003). 'NA' indicates not applicable." %>% 
  str_wrap(width = 100) %>%
  str_split("\\n") %>%
  unlist %>%
  purrr::map_chr(wrap_quotes)

# Write out table with metadata
write_csv_with_meta(
  taxonomic_data_for_si, 
  here::here("results/Table_S1.csv"), 
  header,
  footer)
```

**`r s_table("trait-values")`** `r s_table_cap("trait-values")` (see separate file)

SLA, specific leaf area. Continuous measurements shown as mean, standard deviation (sample size). 'taxon code' is the name used for that taxon in analyses (see Table S1). For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI. 'NA' indicates missing data.

```{r trait_data}
# Set header and footer
header <- c(
  si_table_header,
  s_table_nums("trait-values") %>% wrap_quotes
)

footer <- "SLA, specific leaf area. Continuous measurements shown as mean, standard deviation (sample size). 'taxon code' is the name used for that taxon in analyses (see Table S1). For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI. 'NA' indicates missing data." %>% 
  str_wrap(width = 100) %>%
  str_split("\\n") %>%
  unlist %>%
  purrr::map_chr(wrap_quotes)

# Write out table with metadata
write_csv_with_meta(
  trait_data_for_si, 
  here::here("results/Table_S2.csv"), 
  header,
  footer)
```

**`r s_table("ancova")`** `r s_table_cap("ancova")` (see separate file)

@@

```{r climate-ancova}
# Format table
climate_ancova_results_formatted <-
climate_ancova_results %>% 
  mutate(
    p.value = signif(p.value, 2) %>% format.pval(., digits = 1),
    meansq = round_t(meansq, 2),
    term = case_when(
      term == "habit" ~ "Growth habit",
      term == "el" ~ "Elevation",
      term == "habit:el" ~ "Growth habit × elevation",
      TRUE ~ term
    )
  ) %>%
  select(`Response variable` = variable, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value)

# Set header and footer
header <- c(
  si_table_header,
  s_table_nums("ancova") %>% wrap_quotes
)

footer <- "d.f., degrees of freedom; MS, mean squares." %>% wrap_quotes

# Write out table with metadata
write_csv_with_meta(
  climate_ancova_results_formatted, 
  here::here("results/Table_S3.csv"), 
  header,
  footer)
```

**`r s_table("phylosig-binary-strict")`** `r s_table_cap("phylosig-binary-strict")` (see separate file)

```{r phylosig-binary-strict}
# Format table
phylosig_binary_traits_strict_formatted <-
  phylosig_binary_traits_strict %>%
  mutate(D = round_t(D, 2)) %>%
  mutate(trait = factor(
    trait, 
    levels = c("habit", "gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Epiphytic growth", "Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
  arrange(trait) %>%
  select(Trait = trait, `Number of presences` = num_present, `Number of absences` = num_absent, D,
         P_rnd = prob_random, P_bm = prob_brownian)

# Set header and footer
header <- c(
  si_table_header,
  s_table_nums("phylosig-binary-strict") %>% wrap_quotes
)

footer <- "P_rnd, probability of obtaining the observed value of D under random trait evolution; P_bm, probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion." %>%
  str_wrap(width = 100) %>%
  str_split("\\n") %>%
  unlist %>%
  purrr::map_chr(wrap_quotes)

# Write out table with metadata
write_csv_with_meta(
  phylosig_binary_traits_strict_formatted, 
  here::here("results/Table_S4.csv"), 
  header,
  footer)
```

**`r s_table("corr-evo-strict")`** `r s_table_cap("corr-evo-strict")` (see separate file)

```{r corr-evo-strict}
# Format table
correlated_evo_test_strict_formatted <-
  correlated_evo_test_strict %>%
  mutate(trait = factor(
    trait, 
    levels = c("gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Gemmae", "Glands", "Hairs", "Morphotype"))) %>%
  arrange(trait) %>%
  select(
    Trait = trait,
    `LL (independent model)` = logL_indep,
    `LL (dependent model)` = logL_dep,
    `Likelihood ratio` = likelihood_ratio,
    `P` = pval
  )

# Set header and footer
header <- c(
  si_table_header,
  s_table_nums("corr-evo-strict") %>% wrap_quotes
)

footer <- "LL, log likelihood. Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs. terrestrial). Lower log likelihood for the dependent model indicates correlation with growth habit." %>%
  str_wrap(width = 100) %>%
  str_split("\\n") %>%
  unlist %>%
  purrr::map_chr(wrap_quotes)

# Write out table with metadata
write_csv_with_meta(
  correlated_evo_test_strict_formatted, 
  here::here("results/Table_S5.csv"), 
  header,
  footer)
```

**`r s_table("model-fit")`** `r s_table_cap("model-fit")` (see separate file)

```{r model-fit}
# Format table
best_fit_div_models_moran_formatted <-
best_fit_div_models_moran %>%
  # Drop frond length and width (highly correlated with rhizome dia.)
  filter(resp_var %in% c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna")) %>%
  mutate(resp_var = factor(
        resp_var, 
        levels = c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna"))) %>%
  mutate(
    resp_var = lvls_revalue(
        resp_var, 
        new_levels = c("Richness", "MPDphy", "MNTDphy", 
                       "MPDfunc", "MNTDfunc",
                       "Dissection", "SLA", "Stipe length", "Rhizome diam.", "Pinna no.")) 
  ) %>%
  arrange(resp_var, delta_AICc) %>%
  mutate(
    modname = modname %>%
      str_replace_all("\\+", " \\+ ") %>%
      str_replace_all("\\.by\\.", " x ") %>%
      str_replace_all("habit", "Growth habit") %>%
      str_replace_all("temp_mean", "Mean temp.") %>%
      str_replace_all("temp_sd", "SD temp.") %>%
      str_replace_all("temp_max", "Max. temp.") %>%
      str_replace_all("vpd_mean", "Mean VPD") %>%
      str_replace_all("vpd_sd", "SD VPD") %>%
      str_replace_all("vpd_min", "Min. VPD")
  ) %>% 
  select(
    `Metric` = resp_var,
    `Model` = modname,
    AICc,
    `R2` = r2,
    delta_AICc,
    `Moran's I` = morans_I,
    `P_moran` = morans_I_pval
  )

# Set header and footer
header <- c(
  si_table_header,
  s_table_nums("model-fit") %>% wrap_quotes
)

footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I. Models within delta AICc 3 of the best-fitting model shown. Trait names refer to community-weighted mean values (see Methods). Frond length and width were highly correlated with rhizome diameter and not shown. For details on calculations of MPD and MNTD, see Methods." %>%
  str_wrap(width = 100) %>%
  str_split("\\n") %>%
  unlist %>%
  purrr::map_chr(wrap_quotes)

# Write out table with metadata
write_csv_with_meta(
  best_fit_div_models_moran_formatted, 
  here::here("results/Table_S6.csv"), 
  header,
  footer)

@@

@@

```{r Fig-S1, fig.height = 6, fig.width = 7, fig.cap = ""}
loadd(cwsd_scatterplot, cache = canopy_cache)

cwsd_scatterplot
```

**`r s_figure("cwsd")`** `r s_figure_cap("cwsd")`. Epiphytic communities in green, terrestrial communities in brown. Points randomly jittered along x-axis to reduce overlap. Boxplots show median values (bold lines) by growth habit. Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range. Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; \*\*\* = \pval{ }< 0.001; \* = \pval{ }< 0.05 **(a)**, traits related to plant size **(b)**, traits not related to plant size. Response variable abbreviations: SLA, specific leaf area.

```{r Fig-S2, fig.height = 8, fig.width = 4, fig.cap = ""}
